### 2016-10-04 Paper_v7.tex
Fixed figure numbering ... I think

\subsection*{Combined TGF$-\beta$2 and VEGF-A exposure drives heterogeneity in MCF10A and DLD1 cells}
The EMT model simulations suggested the transcriptional activity of NFATc and Sp1 could be independently tuned to generate a hybrid cell population with both epithelial and mesenchymal characteristics.
To test this hypothesis, we exposed either quiescent epithelial (MCFA10, (Fig. \ref{fg:F5})) or transformed epithelial cells (DLD1, (Fig. \ref{fg:S3})) to combinations of TGF$-\beta$1/2 and/or VEGF-A. As expected, treatment with TGF$-\beta$1/2 (10ng/ml) increased Slug and Vimentin expression, while repressing E-cadherin expression both at the transcript and protein levels in MCF10A (Fig. \ref{fg:F5}A-B) and DLD1 cells (Fig. \ref{fg:S4}C, Fig S3 D,E).
Both MCF10A (Fig. \ref{fg:F5}C) and DLD1 cells (Fig. \ref{fg:S3}E,G) transitioned from quiescent cobblestone morphology to spread spindle shapes, consistent with EMT.
As predicted, we found increased nuclear localization of phosphorylated Sp1 following TGF$-\beta$1/2 stimulation in both MCF10A (Fig. \ref{fg:F5}B,C) and DLD1 cells (Fig. \ref{fg:S3}E,F).
Consistent with model predictions, VEGF-A (50ng/ml) treatment increased the abundance of NFATc1 and E-cadherin at both the transcript and protein level in both MCF10A (Fig. \ref{fg:F5}A) and DLD1 (Fig. \ref{fg:S3}A) cells.
We also found that NFATc1 nuclear localization significantly increased in both MCF10 and DLD1 treated with VEGF-A independently of the abundance of nuclear localized phosphorylated Sp1 levels (Fig. \ref{fg:F5}B,C Fig.S3C,E ).
Interestingly, combining VEGF-A (50ng/ml) with TGF$-\beta$1/2 (10ng/ml) resulted in significantly elevated expression of both E-cadherin and Vimentin at the transcript and protein levels in both MCF10A and DLD1 cells (Fig 5A,B; Fig S3D,E; Fig S4C).
NFATc1 expression increased, while Sp1 expression was similar to the TGF$-\beta$1/2 case alone (Fig. \ref{fg:F5}A-B,  Fig S3D,E; Fig S4C)), supporting their independent regulation.
The expression of Slug, and Vimentin significantly increased, while E-cadherin levels were increased in MCF10A cells (Fig 5A) and maintained at control levels in DLD1 cells (Fig. S3D).y
As further predicted, nuclear co-localization of both NFATc1 and phosphorylated Sp1 were apparent in MCF10A and DLD1 cells treated with both ligands (Fig. \ref{fg:F5}B,C Fig S3E,F).
Taken together, combined VEGF-A and TGF$-\beta$1/2 treatment elicited a hybrid phenotype expressing both mesenchymal and epithelial characteristics in both MCF10A and DLD1 cells.
This phenotype was driven by the transcriptional activity of two key transcription factors, Sp1 and NFATc, which could be modulated independently by TGF$-\beta$1/2 and VEGF-A exposure.

Our phenotypic analysis predicted that NFATc transcriptional activity was critical to maintaining E-cadherin expression in the presence of both VEGF-A and TGF$-\beta$1/2.
We experimentally tested this hypothesis by exposing both MCF10A (Fig. \ref{fg:F5}E,F) and DLD1 cells (Fig. \ref{fg:S4}) to combinations of VEGF-A and TGF$-\beta$1/2 in the presence or absence of VIVIT, a soluble peptide inhibitor of NFATc transcriptional activity \citep{Aramburu:1999if}.
Treatment with VEGF-A (50ng/ml) and VIVIT (10$\mu$M) in MCF10A cells significantly reduced E-cadherin expression compared to VEGF-A alone (Fig 5D,E).
Co-treatment with VIVIT and TGF$-\beta$1/2 did not enhance EMT capacity of MCF10A cells above that of TGF$-\beta$1/2 alone (Fig 5A,B,E).
Likewise, VIVIT in combination with both TGF$-\beta$1/2 and VEGF-A resulted in a loss of E-cadherin gene and protein expression, while Slug and Vimentin levels remained increased (Fig. \ref{fg:F5}D,E ).
Quantitative flow cytometry confirmed these results in both MCF10A (Fig. \ref{fg:F5}F) and DLD1 cells (Fig. S4C).
Both epithelial cell lines initially had high levels of E-cadherin expression, and low vimentin abundance (Q1-99.5\%), but both MCF10A and DLD1 cells shifted from an epithelial to mesenchymal phenotype (Q1-33.4\%, Q4-42.8\%) following TGF$-\beta$1/2 exposure.
As expected, NFATc nuclear localization was repressed with VIVIT treatment regardless of ligand stimulation, while the abundance of nuclear phosphorylated Sp1 increased for both TGF$-\beta$1/2 and TGF$-\beta$1/2 + VIVIT conditions (Fig. 5D,E).
Combined TGF$-\beta$1/2 and VEGF-A increased both Vimentin and E-cadherin expression (Q1-42.1\%, Q2-52.3\%) compared to TGF$-\beta$1/2 alone.
Together, these results demonstrate that NFATc and phosphorylated Sp1 are critical for regulating E-cadherin and Vimentin expression during phenotype heterogeneity in MCF10A and DLD1.

>>>>

\subsection*{Combined TGF$-\beta$2 and VEGF-A exposure drives heterogeneity in MCF10A and DLD1 cells}
The EMT model simulations suggested the transcriptional activity of NFATc and Sp1 could be independently tuned to generate a hybrid cell population with both epithelial and mesenchymal characteristics.
To test this hypothesis, we exposed either quiescent epithelial (MCFA10, Fig. \ref{fg:F5}) or transformed epithelial cells (DLD1, Fig. \ref{fg:S3}) to combinations of TGF$-\beta$1/2 and/or VEGF-A. As expected, TGF$-\beta$1/2 treatment (10ng/ml) increased Slug and Vimentin expression, while repressing E-cadherin expression both at the transcript and protein levels in MCF10A (Fig. \ref{fg:F5}A-B) and DLD1 cells (Fig. \ref{fg:S4}C).
Both MCF10A (Fig. \ref{fg:F5}C) and DLD1 cells (Fig. \ref{fg:S3}E,G) transitioned from quiescent cobblestone morphology to spread spindle shapes, consistent with EMT.
As predicted, we found increased nuclear localization of phosphorylated Sp1 following TGF$-\beta$1/2 stimulation in both MCF10A (Fig. \ref{fg:F5}B,C) and DLD1 cells (Fig. \ref{fg:S3}E,F).
Consistent with model predictions, VEGF-A (50ng/ml) treatment increased the abundance of NFATc1 and E-cadherin at both the transcript and protein level in both MCF10A (Fig. \ref{fg:F5}A) and DLD1 cells (Fig. \ref{fg:S3}).
We also found that NFATc1 nuclear localization significantly increased in both MCF10 (Fig. \ref{fg:F5}B,C) and DLD1 (Fig. \ref{fg:S3}C,E) cells treated with VEGF-A.
Interestingly, combining VEGF-A (50ng/ml) with TGF$-\beta$1/2 (10ng/ml) resulted in significantly elevated expression of both E-cadherin and Vimentin at the transcript and protein levels in both MCF10A (Fig. \ref{fg:F5}A,B) and DLD1 cells (Fig. \ref{fg:S3}-\ref{fg:S4}).
NFATc1 expression increased, while Sp1 expression was similar to the TGF$-\beta$1/2 case alone (Fig. \ref{fg:F5}A-B and Fig. \ref{fg:S3}D,E),
supporting their independent regulation.
The expression of Slug, and Vimentin significantly increased, while E-cadherin levels were increased in MCF10A cells (Fig. \ref{fg:F5}A) and maintained at control levels in DLD1 cells
(Fig. \ref{fg:S3}D).
As predicted, nuclear co-localization of both NFATc1 and phosphorylated Sp1 were apparent in MCF10A (Fig. \ref{fg:F5}B,C) and DLD1 (Fig. \ref{fg:S3}E,F) cells treated with both ligands.
Taken together, combined VEGF-A and TGF$-\beta$1/2 treatment elicited a hybrid phenotype expressing both mesenchymal and epithelial characteristics in both MCF10A and DLD1 cells.
This phenotype was driven by the transcriptional activity of two key transcription factors, Sp1 and NFATc, which could be modulated independently by TGF$-\beta$1/2 and VEGF-A exposure.

Our phenotypic analysis predicted that NFATc transcriptional activity was critical to maintaining E-cadherin expression in the presence of both VEGF-A and TGF$-\beta$1/2.
We experimentally tested this hypothesis by exposing both MCF10A (Fig. \ref{fg:F5}E,F) and DLD1 cells (Fig. \ref{fg:S4}) to combinations of VEGF-A and TGF$-\beta$1/2 in the presence or absence of VIVIT, a soluble peptide inhibitor of NFATc transcriptional activity \citep{Aramburu:1999if}.
Treatment with VEGF-A (50ng/ml) and VIVIT (10$\mu$M) in MCF10A cells significantly reduced E-cadherin expression compared to VEGF-A alone (Fig. \ref{fg:F5}D,E).
Co-treatment with VIVIT and TGF$-\beta$1/2 did not enhance EMT capacity of MCF10A cells above that of TGF$-\beta$1/2 alone (Fig. \ref{fg:F5}A,B,E).
Likewise, VIVIT in combination with both TGF$-\beta$1/2 and VEGF-A resulted in a loss of E-cadherin gene and protein expression, while Slug and Vimentin levels remained increased (Fig. \ref{fg:F5}D,E).
Quantitative flow cytometry confirmed these results in both MCF10A (Fig. \ref{fg:F5}F) and DLD1 cells (Fig. \ref{fg:S4}C).
Both epithelial cell lines initially had high levels of E-cadherin expression, and low Vimentin abundance (Q1-99.5\%), but both MCF10A and DLD1 cells shifted from an epithelial to mesenchymal phenotype (Q1-33.4\%, Q4-42.8\%) following TGF$-\beta$1/2 exposure.
As expected, NFATc nuclear localization was repressed with VIVIT treatment regardless of ligand stimulation, while the abundance of nuclear phosphorylated Sp1 increased for both TGF$-\beta$1/2 and TGF$-\beta$1/2 + VIVIT conditions (Fig. \ref{fg:F5}C,E).
Combined TGF$-\beta$1/2 and VEGF-A increased both Vimentin and E-cadherin expression (Q1-42.1\%, Q2-52.3\%) compared to TGF$-\beta$1/2 alone.
Together, these results demonstrate that NFATc and phosphorylated Sp1 are critical for regulating E-cadherin and Vimentin expression during phenotype heterogeneity in MCF10A and DLD1.
